\documentclass[source/paper/main.tex]{subfiles}
\begin{document}

\section{Model}
\subsection{Baseline model}

Time is discrete and indexed by $t$.
Each time period $t$, the project $i$ faces a latent number of problems $N_{i,t} \sim~ F_{it}(\cdot)$ where $F_{it}$ is some discrete probability mass function. 
Each problem is characterized by states. 
Within period $t$, let $X_{nt}^{(s)} \in \{\ell,o,r,m\}$ denote the state of problem $n$ after stage $s$ of the pull-request process, where $s \in \{0,1,2,3\}$ indexes stages within the period.

Each problem passes through the following stages:
\begin{enumerate}
    \item leads to an opened pull request,
    \item pull request is reviewed,
    \item pull request is merged.
\end{enumerate}

The state space is
\begin{itemize}
    \item $\ell$: latent problem for which no pull request has been opened,
    \item $o$: pull request opened to address the problem,
    \item $r$: opened pull request reviewed,
    \item $m$: reviewed pull request merged.
\end{itemize}
State $m$ is absorbing.

I will characterize the within-period stage transition structure for any problem in project $i$ at time $t$ by 
$$
\pi_{i,t} =
\begin{pmatrix}
1-\pi_{i,t}^o & \pi_{i,t}^o & 0 & 0 \\
0 & 1-\pi_{i,t}^r & \pi_{i,t}^r & 0 \\
0 & 0 & 1-\pi_{i,t}^m & \pi_{i,t}^m \\
0 & 0 & 0 & 1
\end{pmatrix}. 
$$
This characterization invokes Assumptions~\ref{asm:exchangeability} (Problem exchangeability), \ref{asm:independence} (Problem independence), and \ref{asm:memoryless} (Memoryless stage transitions).

Here, $\pi_{i,t}^o$ is the probability that a latent problem leads to an opened pull request, $\pi_{i,t}^r$ is the probability that an opened pull request is reviewed, and $\pi_{i,t}^m$ is the probability that a reviewed pull request is merged. 
The complementary probabilities $1-\pi_{i,t}^o$, $1-\pi_{i,t}^r$, and $1-\pi_{i,t}^m$ should be interpreted broadly. 
For example, a reviewed pull request could be left unmerged because the requested improvements were never implemented or because the reviewer deemed that the pull request's changes would not improve the software.

Hence, in each period $t$, there will be 
\begin{enumerate}
    \item $N_{i,t}^o = N_{i,t} \cdot \pi_{i,t}^o$ opened pull requests,
    \item $N_{i,t}^r = N_{i,t}^o \cdot \pi_{i,t}^r$ reviewed pull requests, and
    \item $N_{i,t}^m = N_{i,t}^r \cdot \pi_{i,t}^m$ merged pull requests.
\end{enumerate}

My goal is to predict how organizational activity changes when a member leaves. To do that, I need a model of how individual actions contribute to organizational outcomes. 

\subsubsection{Opening, review, and merge as person-specific Bernoulli actions}
Each member $j \in \mathcal J$ contributes to organizational activity by opening, reviewing and merging pull requests. 
For a given latent problem, at most one member can open the pull request. 
Any number of members can review an opened pull request.
For any reviewed pull request, at most one member can merge it.

For any latent problem in time period $t$, let $p_{j,t}^o$ denote the probability that member $j$ is the unique member who opens a pull request to address the problem. Hence, I can define the probability that no member opens a pull request,
$$ p_{\emptyset}^o = 1-\sum_{j\in\mathcal J} p_{j,t}^o. $$
which satisfies $ p_{j,t}^o \geq 0 $ and $ \sum_{j\in\mathcal J}p_{j,t}^o \leq 1. $

For each latent problem, the identity of the opening member is a draw from a categorical distribution over $\mathcal{J} \cup \{\emptyset\}$: member $j$ opens the pull request with probability $p_{j,t}^o$, and no member opens it with probability $p_{\emptyset}^o$. Since draws are independent across problems (Assumption~\ref{asm:problem-invariant}), the joint count vector follows a multinomial distribution:
$$ \left( {A_{j,t}^o}_{j\in\mathcal J}, A_{0t}^o \right) \mid N_{i,t} \sim \text{Multinomial}\left(N_{i,t}; {p_{j,t}^o}_{j\in\mathcal J}, p_{\emptyset}^o \right). $$

This additionally invokes Assumption~\ref{asm:problem-invariant} (Problem-invariant member probabilities).
Hence, the probability that a latent problem leads to an opened pull request is
$$ \pi_{i,t}^o = \sum_{j\in\mathcal J}p_{j,t}^o. $$


I model the quantity of pull requests reviewed by member $j$ in period $t$, $A_{j,t}^r$, as $N_{i,t}^o$ draws from a Binomial distribution:
$$A_{j,t}^r \mid N_{i,t}^o \sim \text{Binomial}(N_{i,t}^o,, p_{j,t}^r),$$
where $p_{j,t}^r$ is member $j$'s probability of reviewing a given opened pull request. 
This also invokes Assumptions~\ref{asm:problem-invariant} (Problem-invariant member probabilities) and~\ref{asm:review-independence} (Independent review decisions).
Hence, the probability that an opened pull request is reviewed by at least one member is
$$ \pi_{i,t}^r = 1-\prod_{j\in\mathcal J}(1-p_{j,t}^r). $$


For any reviewed pull request at time $t$, let $p_{j,t}^m$ denote the probability that member $j$ is the unique member who merges it. Hence, I can define the probability that no member merges a pull request,
$$ p_{\emptyset}^m = 1-\sum_{j\in\mathcal J} p_{j,t}^m. $$
which satisfies $ p_{j,t}^m \geq 0 $ and $ \sum_{j\in\mathcal J}p_{j,t}^m \leq 1. $
For each reviewed pull request, the identity of the merging member is a draw from a categorical distribution over $\mathcal{J} \cup \{0\}$: member $j$ merges it with probability $p_{j,t}^m$, and no member merges it with probability $p_{0,t}^m$. By the same independence argument as for opening, the joint count vector follows a multinomial distribution:
$$ \left( {A_{j,t}^m}_{j\in\mathcal J}, A_{0t}^m \right) \mid N_{i,t}^r \sim \text{Multinomial}\left(N_{i,t}^r; {p_{j,t}^m}_{j\in\mathcal J}, p_{0,t}^m \right). $$
Hence, the probability that a reviewed pull request is merged is $$ \pi_{i,t}^m = \sum_{j\in\mathcal J}p_{j,t}^m. $$
This also invokes Assumption~\ref{asm:problem-invariant} (Problem-invariant member probabilities).


\subsection{Prediction}
My goal is to predict, after a member $j^*$ departs in time period $t$, within an organization, the next-period $(t+1)$ quantity of
\begin{enumerate}
    \item pull requests opened ($N_{i,t+1}^o$),
    \item pull requests reviewed ($N_{i,t+1}^r$), and
    \item pull requests merged ($N_{i,t+1}^m$)
\end{enumerate}
I do that by fitting a model of the latent problem flow and member-level probabilities on organizational pre-period data, and then using the fitted model to predict counterfactual next-period activity after removing member $j^*$.

I impose two modeling choices to identify and estimate the latent problem flow:
\begin{enumerate}
    \item \textbf{Normalization of unobserved fraction}: $p_{\emptyset,t}^o = 0$ during the pre-periods, so every latent problem generates an opened pull request. This directly identifies the latent problem count from observed data, since $N_{i,t} = N_{i,t}^o$.
    \item \textbf{Time-invariant latent flow}: $F_{i,t} = F_i$, so pre-period opening counts $\{N_{i,t}^o\}_{t<0}$ are i.i.d.\ draws from $F_i$, which I estimate by maximum likelihood.
\end{enumerate}

I evaluate three fitting approaches, each applied organization-by-organization on pre-period opening counts by maximum likelihood:
\begin{enumerate}
    \item \textbf{Poisson/ZIP}: fit both Poisson and ZIP for each organization; select between them using AIC.
    \item \textbf{Negative binomial}: fit a negative binomial for each organization.
    \item \textbf{Best of all three}: fit all three candidates and select the best-fitting distribution for each organization using AIC.
\end{enumerate}
Since pre-period lengths are the same across organizations, AIC is directly comparable across organizations and approaches without normalization. For each approach, I report the cross-organization distribution of AIC. To compare approaches, I report the cross-organization distribution of pairwise AIC differences (e.g., approach 1 minus approach 2 per organization), with the median AIC difference as a summary statistic. I also store the empirical versus fitted PMF for each organization as an auxiliary visual check.

% Question for Jesse: Since pre-period lengths are the same across organizations, AIC is directly comparable both within and across organizations. Does it make more sense to select a distributional family separately for each organization (approaches 3), or to impose a common family across all organizations (approach 1, 2)? Is the majority-vote rule --- adopt the family that wins for more than half of organizations --- a reasonable way to select a common family?}

I impose three additional modeling choices governing member probabilities:
\begin{enumerate}
    \item \textbf{Stationary member probabilities}: $p_{j,t}^s = p_j^s$ for all $t$ and all $s$. Under stationarity, pre-period observations pool cleanly across periods for estimation, and the counterfactual after departure is the model evaluated at the same probabilities with member $j^*$ removed.
    \item \textbf{Fixed membership}: the membership set $\mathcal{J}$ is fixed across periods, with no entry or exit other than the focal departure $j^*$.
\end{enumerate}
Note that the remaining members' probabilities are unchanged after the departure. This is follows directly from the Assumption~\ref{asm:problem-invariant} (Problem-invariant member probabilities).

I use one estimation approach for member probabilities:
\begin{enumerate}
    \item \textbf{Pooled pre-period shares}: all member probabilities are identified by pooled pre-period shares, requiring no numerical optimization. For opening and merging, the MLE under a stationary multinomial is $\hat{p}_j^o = \frac{\sum_t A_{j,t}^o}{\sum_t N_{i,t}^o}$ and $\hat{p}_j^m = \frac{\sum_t A_{j,t}^m}{\sum_t N_{i,t}^r}$. For reviewing, the Binomial MLE is $\hat{p}_j^r = \frac{\sum_t A_{j,t}^r}{\sum_t N_{i,t}^o}$.
\end{enumerate}
For opening and merging, I evaluate fit via AIC. For each organization, I compute the multinomial log-likelihood of each pre-period's member counts under the pooled fitted shares $\hat{p}_j^o$ (respectively $\hat{p}_j^m$), sum across pre-periods to obtain $\ell_i^o = \sum_t \log p(A_{j,t}^o \mid N_{i,t}^o, \hat{p}_j^o)$, and set $\text{AIC}_i^o = -2\ell_i^o + 2(|\mathcal{J}|-1)$. This yields one AIC per organization. I report the cross-organization distribution of AIC, with the median as a summary statistic. I also store the empirical versus fitted member share distribution for each organization and period as an auxiliary visual check.

For reviewing, fit evaluation has two components. First, I compute AIC analogously: for each organization, I sum the per-member Binomial log-likelihoods across members and pre-periods, $\ell_i^r = \sum_t \sum_j \log p(A_{j,t}^r \mid N_{i,t}^o, \hat{p}_j^r)$, and set $\text{AIC}_i^r = -2\ell_i^r + 2|\mathcal{J}|$, yielding one AIC per organization. I report the cross-organization distribution of AIC with the median as a summary statistic. Second, I check whether the implied aggregate review probability $\hat{\pi}^r = 1-\prod_{j\in\mathcal{J}}(1-\hat{p}_j^r)$ matches the observed per-period review rate $\tilde{\pi}_t^r = N_{i,t}^r / N_{i,t}^o$, reporting the cross-organization distribution of the mean squared deviation between $\hat{\pi}^r$ and $\tilde{\pi}_t^r$ across pre-periods, with the median as a summary statistic.


% next steps
% 1. Reread the full doc and see where I should amke improvements
% 2. Formalize the comments for Jesse
% 3. Write out the different sources of prediction error, and how I can figure out how important each component is?

%- ASK JESSE: is how i'm presenting things here clear?
%- ASK JESSE: how do I know when good is good enough? I won't get perfect prediction performance. I don't even know if I"ll be distributionally good enough. 



%MODELLING choices I will not make now but might be good to incorporate
%- model entry/exit from the org 
%- Is there a way to penalize big p's?
%- find ways to loosen assumptions - I feel like I have 5 structural ones + the modelling ones above.  - ACTUALLY UPON FURTHER REFLECTION, I THINK I SHOULD THINK ABOUT HOW I"M GOING TO VARY THE MODEL BASED ON WHAT I"M FAILING TO PREDICT WELL. THE TRICKY THING HERE IS THAT I ONLY HAVE ONE OBSERVATION FOR "HOW WELL I"M PREDICTING". SO I HAVE TO THINK ABOUT HOW DISTRIBUTIONALLY ACROSS ORGS MODELLING CHOICES IMPROVE PREDICTIVE PERFORMANCE??? 


I use the following notation for fitted and predicted quantities. Let $\hat{p}_j^s$ denote the estimated member probability for member $j$ at stage $s \in \{o, r, m\}$, computed from pooled pre-period shares as described above. Let $\hat{F}_i$ denote the fitted latent flow distribution for organization $i$, and let $\hat{N}_i$ denote the corresponding point estimate (the mean of $\hat{F}_i$ draws). Let $\hat{\pi}_i^{s,(-j^*)}$ denote the counterfactual stage-$s$ completion probability after removing member $j^*$, constructed from the $\hat{p}_j^s$. Finally, let $\hat{N}_{i,t+1}^s$ denote the predicted count of stage-$s$ activity in the next period $t+1$.

Under this model, the counterfactual stage-completion probabilities and implied counts after removing member $j^*$ are as follows.

\begin{enumerate}
    \item $$ \hat{\pi}_i^{o,(-j^*)} = \sum_{j\in\mathcal J\setminus\{j^*\}}\hat{p}_j^o, $$ so the counterfactual number of pull requests opened is $$ \hat{N}_{i,t+1}^o = \hat{N}_i\,\hat{\pi}_i^{o,(-j^*)}. $$
    \item The counterfactual review probability is
    $$ \hat{\pi}_i^{r,(-j^*)} = 1-\prod_{j\in\mathcal J\setminus\{j^*\}}(1-\hat{p}_j^r), $$ so the counterfactual number of pull requests reviewed is $$ \hat{N}_{i,t+1}^r = \hat{N}_i\,\hat{\pi}_i^{o,(-j^*)}\hat{\pi}_i^{r,(-j^*)}. $$
    \item The counterfactual merge probability is $$ \hat{\pi}_i^{m,(-j^*)} = \sum_{j\in\mathcal J\setminus\{j^*\}}\hat{p}_j^m, $$ so the counterfactual number of pull requests merged is $$ \hat{N}_{i,t+1}^m = \hat{N}_i\,\hat{\pi}_i^{o,(-j^*)}\hat{\pi}_i^{r,(-j^*)}\hat{\pi}_i^{m,(-j^*)}. $$
\end{enumerate}


\subsubsection{Putting everything together}


\label{sec:stage_decomposition_baseline}

The stage-completion probabilities defined above together imply
$$
E[Q_t^m \mid N_{i,t}]
=
N_{i,t} \pi_{i,t}^o \pi_{i,t}^r \pi_{i,t}^m.
$$

Now define member $j$'s marginal contribution at each stage as the change in the stage-completion probability when $j$ is removed:
$$
\Delta_o^{(j)}
\equiv
\pi_{i,t}^o - \pi_{i,t}^o^{(-j)}
=
p_{j,t}^{o}\prod_{j\in\mathcal J\setminus\{j\}}(1-p_{j,t}^{o}),
$$
$$
\Delta_r^{(j)}
\equiv
\pi_{i,t}^r - \pi_{i,t}^r^{(-j)}
=
p_{j,t}^{r}\prod_{j\in\mathcal J\setminus\{j\}}(1-p_{j,t}^{r}),
$$
$$
\Delta_m^{(j)}
\equiv
\pi_{i,t}^m - \pi_{i,t}^m^{(-j)}
=
p_{j,t}^{m}\prod_{j\in\mathcal J\setminus\{j\}}(1-p_{j,t}^{m}).
$$

Suppose member $j$ leaves before period $t+1$ and is not replaced, and suppose $N_{i,t}=N_{t+1}=N$. Then
$$
E[Q_{t+1}^m \mid N]
=
N \pi_{i,t}^o^{(-j)} \pi_{i,t}^r^{(-j)} \pi_{i,t}^m^{(-j)}
=
N(\pi_{i,t}^o-\Delta_o^{(j)})(\pi_{i,t}^r-\Delta_r^{(j)})(\pi_{i,t}^m-\Delta_m^{(j)}).
$$
Therefore,
$$
E[Q_{t+1}^m \mid N] - E[Q_t^m \mid N]
=
-N\Big[
\Delta_o^{(j)}(\pi_{i,t}^r-\Delta_r^{(j)})(\pi_{i,t}^m-\Delta_m^{(j)})
+\Delta_r^{(j)}\pi_{i,t}^o(\pi_{i,t}^m-\Delta_m^{(j)})
+\Delta_m^{(j)}\pi_{i,t}^o \pi_{i,t}^r
\Big].
$$

The three terms correspond to member $j$'s marginal contribution at the opening, review, and merge stages:
\begin{enumerate}
    \item $-N\,\Delta_o^{(j)}(\pi_{i,t}^r-\Delta_r^{(j)})(\pi_{i,t}^m-\Delta_m^{(j)})$ captures the loss from problems that would have been opened because member $j$ acts at the opening stage, and that would then have been reviewed and merged without member $j$'s involvement at those later stages.

    \item $-N\,\pi_{i,t}^o\Delta_r^{(j)}(\pi_{i,t}^m-\Delta_m^{(j)})$ captures the loss from pull requests that would have been opened (by anyone), then reviewed because member $j$ acts at the review stage, and then merged without member $j$'s involvement at the merge stage.

    \item $-N\,\pi_{i,t}^o \pi_{i,t}^r \Delta_m^{(j)}$ captures the loss from pull requests that would have been opened and reviewed (by anyone), and then merged because member $j$ acts at the merge stage.
\end{enumerate}

\subsection{Structural assumptions}

\begin{assumption}[Problem exchangeability]
\label{asm:exchangeability}
Problems are indistinguishable: for any two problems $n$ and $n'$ and any states $x,y \in \{\ell,o,r,m\}$,
$$
\Pr(X_{n,t}^{(s+1)}=y \mid X_{n,t}^{(s)}=x)
=
\Pr(X_{n',t}^{(s+1)}=y \mid X_{n',t}^{(s)}=x).
$$
Conditional on being in the same state, any two problems have the same probability of advancing to the next stage.
\end{assumption}

\begin{assumption}[Problem independence]
\label{asm:independence}
Conditional on current states, problems evolve independently of one another:
$$
\Pr(X_{1,t}^{(s+1)}=x_1,\dots,X_{N_{i,t},t}^{(s+1)}=x_{N_{i,t}}\mid X_{1,t}^{(s)},\dots,X_{N_{i,t},t}^{(s)})
=
\prod_{n=1}^{N_{i,t}}\Pr(X_{n,t}^{(s+1)}=x_n\mid X_{n,t}^{(s)}).
$$
\end{assumption}

\begin{assumption}[Memoryless stage transitions]
\label{asm:memoryless}
A problem's next stage depends only on its current stage, not its history:
$$
\Pr(X_{n,t}^{(s+1)}=y \mid X_{n,t}^{(s)},X_{n,t}^{(s-1)},\dots)
=
\Pr(X_{n,t}^{(s+1)}=y \mid X_{n,t}^{(s)}).
$$
\end{assumption}

\begin{assumption}[Problem-invariant member probabilities]
\label{asm:problem-invariant}
A member's probability of acting does not vary across problems within a period --- $p_{j,t}^s$ is the same regardless of which particular problem is being acted on. This is distinct from Assumption~\ref{asm:exchangeability}, which describes the distribution problems are drawn from: even if problems differed objectively, this rules out members selectively targeting particular problems.
\end{assumption}

Assumptions \ref{asm:exchangeability}--\ref{asm:problem-invariant} imply that every problem follows the same stage-transition matrix $P$ so that, for every problem $n$,
$$
\Pr(X_{n,t}^{(s+1)}=y \mid X_{n,t}^{(s)}=x)=P_{xy}.
$$

\begin{assumption}[Independent review decisions]
\label{asm:review-independence}
Conditional on an opened pull request, each member $j$ independently decides whether to review it. Whether member $j$ reviews a pull request does not affect whether member $j'$ does so. Since each such decision is a Bernoulli trial, independence across members implies $A_{j,t}^r \mid N_{i,t}^o \sim \text{Binomial}(N_{i,t}^o, p_{j,t}^r)$.
\end{assumption}

\subsection{Measurement and empirical implementation}

This subsection lays out how the model is taken to data. We separate two roles for organizations. \emph{Control organizations} (never-treated) are used to fit the four counterfactual inputs: an estimation sample for member-level shares, a contributor partition, a latent-problem model, and a latent flow $\hat N$. \emph{Treated organizations} (those experiencing a member departure) are used to evaluate predictions. Throughout, we distinguish member-level shares $p_i^s$ from aggregate stage probabilities $\pi^s$: the latter is the model-implied probability that a stage completes, built from the former via the latent-problem model.

\subsubsection{Error metrics and notation}
\label{sec:error_metrics}

This subsubsection collects the generic error-measurement primitives used in the next two subsubsections (fitting on controls and evaluation on treated organizations). The presentation here is purely mathematical: every input is a generic predicted or observed object indexed by an organization $u$, post-event-time $k=0,\ldots,K$, and---where applicable---an outcome $d\in\mathcal D$. The pre-event window is $k=-L,\ldots,-1$, with per-period count $Q_{u,k}^d$ and pre-event mean
$$
\bar Q_{u,\mathrm{pre}}^d
=
\frac{1}{L}\sum_{k=-L}^{-1}Q_{u,k}^d.
$$

\paragraph{Scaled RMSE.} For a generic predicted trajectory $\hat Z_{u,k}^d$ and target $Z_{u,k}^d$ over $k=0,\ldots,K$,
$$
\mathcal R_u^d(\hat Z,Z;\rho)
=
\left[
\frac{1}{K+1}\sum_{k=0}^{K}
\bigl(\rho_u^d(\hat Z_{u,k}^d-Z_{u,k}^d)\bigr)^2
\right]^{1/2}.
$$
For count-valued outcomes, set $\rho_u^d=1/\bar Q_{u,\mathrm{pre}}^d$, so the metric measures error in pre-event count units. For probability- or share-valued outcomes, set $\rho_u^d=1$.

\paragraph{Aggregate weighted error.} Combine across outcomes $d\in\mathcal D$:
$$
\mathcal R_u^{\mathrm{agg}}
=
\left[
\sum_{d\in\mathcal D}w_u^d\bigl(\mathcal R_u^d\bigr)^2
\right]^{1/2}.
$$
We report results under both pre- and post-event activity weights:
$$
w_{u,\mathrm{pre}}^d
=
\frac{\sum_{k=-L}^{-1}Q_{u,k}^d}
{\sum_{d'\in\mathcal D}\sum_{k=-L}^{-1}Q_{u,k}^{d'}},
\qquad
w_{u,\mathrm{post}}^d
=
\frac{\sum_{k=0}^{K}Q_{u,k}^d}
{\sum_{d'\in\mathcal D}\sum_{k=0}^{K}Q_{u,k}^{d'}}.
$$

\paragraph{Normalized mean error.} For directional bias (signed),
$$
\mathrm{NME}_u^d
=
\frac{1}{K+1}\sum_{k=0}^{K}\rho_u^d(\hat Z_{u,k}^d-Z_{u,k}^d).
$$

\paragraph{Share error over a partition.} A second primitive measures discrepancy between a fitted and observed share vector over a finite partition. Let $\mathcal G$ be a finite cell set with $|\mathcal G|$ cells, $\hat P=(\hat P_g)_{g\in\mathcal G}$ a fitted share vector that sums to a constant across $\mathcal G$, and $\tilde P_{g,k}$ an observed share trajectory over the same cells across $k=0,\ldots,K$. Define
$$
\mathcal R^p(\hat P,\tilde P;\mathcal G)
=
\left[
\frac{1}{K+1}\sum_{k=0}^{K}\sum_{g\in\mathcal G}
\bigl(\hat P_g-\tilde P_{g,k}\bigr)^2
\right]^{1/2},
$$
with per-cell variant
$$
\bar{\mathcal R}^p(\hat P,\tilde P;\mathcal G)
=
\left[
\frac{1}{(K+1)|\mathcal G|}\sum_{k=0}^{K}\sum_{g\in\mathcal G}
\bigl(\hat P_g-\tilde P_{g,k}\bigr)^2
\right]^{1/2}.
$$
Dividing by $|\mathcal G|$ rescales the per-cell version so its magnitude is comparable across partitions of different granularity.

\subsubsection{Fitting counterfactual inputs}
\label{sec:fitting}

The model's counterfactual predictions require four fitted objects: (i) an estimation sample for member-level shares $\hat p_i^s$, (ii) a contributor partition $\mathcal G_{u,c}^s$ that determines which contributors are tracked individually and which are pooled, (iii) a latent-problem model that aggregates member-level shares to stage-level $\hat\pi_u^s$, and (iv) the latent problem flow $\hat N_u$. Each is fit on never-treated organizations using a pseudo-departure date.

Fix a stage $s\in\{o,r,m\}$. Let $A_{u,i,k}^s$ denote member $i$'s stage-$s$ actions at organization $u$ in period $k$, with $A_{u,k}^s\equiv\sum_{i\in\mathcal I_u}A_{u,i,k}^s$. Table~\ref{tab:model_fit} lists the candidate specifications for each object; selection among candidates is by the control-organization fit metrics defined below.

\begin{table}[h]
\caption{Candidate specifications for the four fitted objects}
\label{tab:model_fit}
\centering
\begin{tabular}{p{3.4cm}p{10cm}}
\toprule
Object & Candidates \\
\midrule
\textbf{Estimation sample for $\hat p_i^s$} &
Pooled across $L$ pre-periods: $\hat p_{u,i}^{s,\mathrm{pool}}=\dfrac{\sum_{k=-L}^{-1}A_{u,i,k}^s}{\sum_{k=-L}^{-1}A_{u,k}^s}$.\newline
Last pre-period: $\hat p_{u,i}^{s,\mathrm{last}}=A_{u,i,-1}^s/A_{u,-1}^s$. \\[10pt]
\textbf{Contributor partition $\mathcal G_{u,c}^s$} &
All contributors (one cell per member); top 5 across all activity (top-5 individuals plus an ``other'' cell); union of stage-specific top-5 groups; smallest top-$K$ covering 80\% of all activity; union of stage-specific top-$K$ groups covering 80\% in each stage. Each cell is an individual contributor or the ``other'' aggregate. \\[10pt]
\textbf{Latent-problem model} &
Share-normalized opening: $\hat\pi_u^o=\sum_{i\in\mathcal I_u}\hat p_{u,i}^o=1$, with counterfactual $\hat\pi_u^{o,(-j)}=1-\hat p_{u,j}^o$.\newline
Independent-Bernoulli aggregation for review and merge: $\hat\pi_u^{s,(-j)}=1-\prod_{i\neq j}(1-\hat p_{u,i}^s)$, $s\in\{r,m\}$. \\[10pt]
\textbf{Latent flow $\hat N_u$} &
Pre-period average: $\hat N_u^{\mathrm{avg}}=\frac{1}{L}\sum_{k=-L}^{-1}Q_{u,k}^o$.\newline
Last pre-period: $\hat N_u^{\mathrm{last}}=Q_{u,-1}^o$. \\
\bottomrule
\end{tabular}
\end{table}

\paragraph{Modeling latent problems.} Under the share-normalized opening model, $\hat\pi_u^o=1$ in every pre-period, so observed pre-period openings $Q_{u,k}^o$ identify $N_{u,k}$ directly. The identification issue arises only after removing $j$, when $\hat\pi_u^{o,(-j)}<1$ and observed openings reflect the joint product $N_u\,\hat\pi_u^{o,(-j)}$. For review and merge stages, $\hat p_i^r$ and $\hat p_i^m$ describe the conditional distribution of who acts given that someone does; the independent-Bernoulli aggregation translates these into stage-completion probabilities.

\paragraph{Calibration on control organizations.} Index a candidate estimation sample by $a\in\{\mathrm{pool},\mathrm{last}\}$ and a candidate contributor partition by $c$. For each control organization with a pseudo-departure date $k=0$, define observed post-period cell shares
$$
\tilde p_{u,g,k}^{s}(c)
=
\frac{A_{u,g,k}^{s}(c)}{A_{u,k}^{s}},
\qquad
A_{u,g,k}^{s}(c)
=
\sum_{i\in g}A_{u,i,k}^{s}.
$$
The fit error for member-level shares at stage $s$ is the partition-share error from \S\ref{sec:error_metrics} applied to fitted vs.\ observed cell shares:
$$
\mathcal R_u^{p,s}(a,c)
=
\mathcal R^p\bigl(\hat p_{u,\cdot}^{s}(a,c),\,\tilde p_{u,\cdot,k}^{s}(c);\,\mathcal G_{u,c}^s\bigr),
$$
with per-cell variant $\bar{\mathcal R}_u^{p,s}(a,c)=\bar{\mathcal R}^p(\hat p_{u,\cdot}^{s}(a,c),\,\tilde p_{u,\cdot,k}^{s}(c);\,\mathcal G_{u,c}^s)$. We aggregate across stages using both pre- and post-event activity weights based on $A_{u,k}^s$:
$$
\mathcal R_u^{p,\mathrm{agg}}(a,c)
=
\left[
\sum_{s\in\{o,r,m\}}w_u^s\bigl(\mathcal R_u^{p,s}(a,c)\bigr)^2
\right]^{1/2},
\qquad
w_{u,\mathrm{pre}}^s
=
\frac{\sum_{k=-L}^{-1}A_{u,k}^s}{\sum_{s'}\sum_{k=-L}^{-1}A_{u,k}^{s'}},
$$
with analogous post-event weights. Selection over $(a,c)$ minimizes $\mathcal R_u^{p,\mathrm{agg}}$ averaged across the control set.

For latent flow $\hat N_u$, selection uses first-post-period opening error. With control set $\mathcal C$,
$$
\mathcal R^{N}
=
\left[
\frac{1}{|\mathcal C|}\sum_{u\in\mathcal C}
\left(\frac{\hat N_u-Q_{u,0}^o}{\bar Q_{u,\mathrm{pre}}^o}\right)^2
\right]^{1/2}.
$$
We also report the unnormalized RMSE alongside, and overlay the cross-organization distribution of $\hat N$ over event time as a visual check that the chosen rule is broadly consistent with the data.

\paragraph{Plots.} For each candidate specification in Table~\ref{tab:model_fit}, plot the cross-organization distribution of the relevant error metric: $\mathcal R_u^{p,s}$ per stage and the aggregate $\mathcal R_u^{p,\mathrm{agg}}$ for the share-fit candidates, and $(\hat N_u-Q_{u,0}^o)/\bar Q_{u,\mathrm{pre}}^o$ for the $\hat N$ candidates.

\todo[inline]{Joint optimization. Each of the four objects is currently tuned separately. A joint alternative would minimize a single objective on controls that combines targets across objects --- candidates include member-level $p_i^s$, aggregate $\pi^s$, count outcomes $Q^s$, transition rates, opening flow $N\pi_o$, or a weighted combination. More rigorous, but less transparent and harder to diagnose; pursue as an extension if time permits.}

\subsubsection{Evaluating departure predictions}

Given the fitted objects from \S\ref{sec:fitting}, the model predicts post-departure activity at a treated organization $u$ for $k\geq 0$:
$$
\hat Q_{u,k}^{o,(-j)}
=
\hat N_u\,\hat\pi_u^{o,(-j)},
$$
$$
\hat Q_{u,k}^{r,(-j)}
=
\hat N_u\,\hat\pi_u^{o,(-j)}\,\hat\pi_u^{r,(-j)},
$$
$$
\hat Q_{u,k}^{m,(-j)}
=
\hat N_u\,\hat\pi_u^{o,(-j)}\,\hat\pi_u^{r,(-j)}\,\hat\pi_u^{m,(-j)}.
$$
Evaluation proceeds in five steps.

\paragraph{1. Count prediction error.} Report $\mathcal R_u^d$ from \S\ref{sec:error_metrics} for each $d\in\{o,r,m\}$ on treated organizations, plus the aggregate $\mathcal R_u^{\mathrm{agg}}$ under both pre- and post-event weights. Plot the cross-organization distribution of each $\mathcal R_u^d$ and of $\mathcal R_u^{\mathrm{agg}}$ across treated organizations.

\paragraph{2. Treated individual-share diagnostic.} The fit-on-controls diagnostic from \S\ref{sec:fitting} measures how well $\hat p_i^s$ tracks future shares on never-treated organizations. We also want to know how well it tracks shares on the treated organizations themselves, restricting to remaining members. Let $\mathcal G_{u,c}^{s,(-j)}$ denote the contributor cells after removing departed member $j$. Renormalize the fitted treated pre-period composition over remaining cells, and form observed post-period composition over the same restricted set:
$$
\hat\omega_{u,g}^{s,(-j)}(a,c)
=
\frac{\hat p_{u,g}^{s}(a,c)}{\sum_{h\in\mathcal G_{u,c}^{s,(-j)}}\hat p_{u,h}^{s}(a,c)},
\qquad
\tilde\omega_{u,g,k}^{s,\mathrm{post}}(c)
=
\frac{A_{u,g,k}^{s}(c)}{\sum_{h\in\mathcal G_{u,c}^{s,(-j)}}A_{u,h,k}^{s}(c)}.
$$
Apply the partition-share error from \S\ref{sec:error_metrics}:
$$
\mathcal R_{u,\mathrm{treated}}^{p,s}(a,c)
=
\mathcal R^p\bigl(\hat\omega_{u,\cdot}^{s,(-j)}(a,c),\,\tilde\omega_{u,\cdot,k}^{s,\mathrm{post}}(c);\,\mathcal G_{u,c}^{s,(-j)}\bigr),
$$
with per-cell variant $\bar{\mathcal R}_{u,\mathrm{treated}}^{p,s}$ analogous. This individual-level diagnostic is reported separately from the outcome-relevant aggregate $\pi^s$ error: it asks whether composition among remaining members is stable, not whether stage-completion probabilities themselves are correct. Plot the cross-organization distribution per stage.

\paragraph{3. Implied aggregate stage rates.} Define implied (data-side) rates
$$
\tilde\pi_{u,k}^{o\mid\hat N}
=
\frac{Q_{u,k}^o}{\hat N_u},
\qquad
\tilde\pi_{u,k}^{r}
=
\frac{Q_{u,k}^r}{Q_{u,k}^o},
\qquad
\tilde\pi_{u,k}^{m}
=
\frac{Q_{u,k}^m}{Q_{u,k}^r}.
$$
Note that $\tilde\pi_{u,k}^{o\mid\hat N}$ is \emph{not} observed true $\pi_{i,t}^o$; it is the opening probability implied by observed openings under the chosen $\hat N_u$. Values exceeding $1$ flag $\hat N_u$ as too small for observed openings. Plot each $\tilde\pi_{u,k}^{s}$ over event time $k$ across treated organizations and overlay the model's $\hat\pi_u^{s,(-j)}$ as a reference.

\paragraph{4. Progressive oracle attribution.} To localize where the model loses accuracy, hold $\hat N_u$ fixed and substitute oracle (implied) stage rates one stage at a time. The substitutions are:
$$
\hat Q_{\mathrm{oracle}\,o}^{o}=Q^o,
$$
$$
\hat Q_{\mathrm{oracle}\,o}^{r}
=
\hat N_u\,\tilde\pi_{u,k}^{o\mid\hat N}\,\hat\pi_u^{r,(-j)},
\qquad
\hat Q_{\mathrm{oracle}\,o,r}^{r}=Q^r,
$$
$$
\hat Q_{\mathrm{oracle}\,o}^{m}
=
\hat N_u\,\tilde\pi_{u,k}^{o\mid\hat N}\,\hat\pi_u^{r,(-j)}\,\hat\pi_u^{m,(-j)},
$$
$$
\hat Q_{\mathrm{oracle}\,o,r}^{m}
=
\hat N_u\,\tilde\pi_{u,k}^{o\mid\hat N}\,\tilde\pi_{u,k}^{r}\,\hat\pi_u^{m,(-j)},
\qquad
\hat Q_{\mathrm{oracle}\,o,r,m}^{m}=Q^m.
$$
Re-evaluate $\mathcal R_u^d$ and $\mathcal R_u^{\mathrm{agg}}$ at each oracle level. The drop in error attributable to each substitution localizes the dominant misfit: a large drop from feeding in $\tilde\pi^{o\mid\hat N}$ implicates $\hat N_u$ (since under the share-normalized opening model the only way the opening prediction can be wrong is via $\hat N_u$, given $\hat\pi_u^{o,(-j)}=1-\hat p_{u,j}^o$ is identified by pre-period shares); a large drop from $\tilde\pi^{r}$ implicates the review-stage independent-Bernoulli aggregation; and so on. Plot the cross-organization distribution of $\mathcal R_u^{\mathrm{agg}}$ at each oracle level.

\paragraph{5. Stage decomposition.} The predicted decline in merges decomposes into opening, review, and merge channels (\S\ref{sec:stage_decomposition_baseline}). Compare actual stage quantities $Q_{u,k}^o,Q_{u,k}^r,Q_{u,k}^m$ and stage failures $Q_{u,k}^o-Q_{u,k}^r$ and $Q_{u,k}^r-Q_{u,k}^m$ to predicted counterparts using $\mathcal R_u^d$ from \S\ref{sec:error_metrics}. The failure-quantity comparison is a stricter test than the stage-level counts: the model must correctly predict not only how many pull requests reach each stage but also where in the funnel they drop out. Plot the cross-organization distribution of $\mathcal R_u^d$ across the six quantities.

\subsection{Extensions}

\begin{enumerate}
    \item \textbf{Relax Assumption 1.} Problems need not come from an identical distribution. For example, the identity of the pull request opener may affect the probability that the pull request is eventually merged.

    \item \textbf{Relax Assumption 2.} Problems need not evolve independently. For example, one pull request being merged may increase the likelihood that followup problems are opened or closed.

    \item \textbf{Relax Assumption 4.} Member-specific probabilities may vary over time. For example, one could model $\pi_{i,t}^{o}$, $\pi_{i,t}^{r}$, and $\pi_{i,t}^{m}$ as draws from member-specific distributions, although this may require more data than are currently available.

    \item \textbf{Relax Assumption 5.} Members' actions within a stage need not be independent. For example, reviewers may coordinate so that one person's review makes another person's review less likely, or a member's departure may induce other members to increase their probabilities of acting. This extension would replace the independent-Bernoulli benchmark with a model that allows for coordination and/or organizational adaptation.
\end{enumerate}

\end{document}
